When Subgraph Isomorphism is Really Hard, and Why This Matters for Graph Databases

### 2018-02-14 paper/paper.tex
Minor tweakery

Improved search orderings for VF2-style algorithms also appear in recent work by
\citet{DBLP:conf/gbrpr/CarlettiFV15} and \citet{o:Carletti16}, cumulating in VF3
\citep{DBLP:conf/gbrpr/CarlettiFSV17,o:CarlettiFSV17}, and by \citet{DBLP:conf/wise/ShenZ17}. We
have seen in \cref{section:labelled} that such an approach could give improvements, but will not
bring an algorithm that does not use domains close to the performance of one that does on harder
instances.  We thus believe that a significant amount of effort in designing algorithms for subgraph
isomorphism has been misdirected due to a lack of familiarity with the best existing algorithms for
subgraph isomorphism on one hand, and with broadly-applicable lessons from the constraint
programming literature on the other.

We now move our focus from the importance of using state-of-the-art techniques for algorithm design,
to the pitfalls of choosing unrealistically easy instances to benchmark our algorithms.
This was stressed by
\citet{DBLP:journals/heuristics/Gent98}:

@@

This lesson has not been well heeded. For example, \citet{DBLP:journals/tcbb/BonniciG17}
recently compared various non-constraint-based approaches to LAD using benchmark sets whose
difficulties are measured between microseconds and milliseconds. Their results suggest that LAD is
an order of magnitude slower than other approaches. This should not surprise us: the initialisation
costs of LAD are non-trivial due to it constructing domains, calculating neighbourhood degree
sequences, and performing all-different filtering. However, the risks of occasionally encountering
exponential performance from weaker algorithms on easy instances should now be clear.

@@

Unfortunately, we saw that this aspect of VF2's performance has had practical consequences, with the
most notable being the misdesign of graph database systems. By not trying even the simplest constraint
programming techniques from the literature, and disregarding all that is known about the empirical
hardness of \NP-complete problems, members of the graph databases community have reached the mistaken
conclusion that extensive research into supporting techniques is important, whereas really what
they are doing is working around some but not all of the defects in their chosen algorithms for subgraph
isomorphism. With this new understanding of what does and does not make subgraph
isomorphism hard, it is time for a radical rethink of how graph database systems work.

>>>>

Improved search orderings for VF2-style algorithms also appear in recent work by
\citet{DBLP:conf/gbrpr/CarlettiFV15} and \citet{o:Carletti16}, cumulating in VF3
\citep{DBLP:conf/gbrpr/CarlettiFSV17,o:CarlettiFSV17}, and by \citet{DBLP:conf/wise/ShenZ17}. We
have seen in \cref{section:labelled} that such an approach could give improvements, but will not
bring an algorithm that does not use domains close to the performance of one that does on harder
instances.  We thus believe that a significant amount of effort in designing algorithms for and
systems involving subgraph isomorphism has been misdirected due to a lack of familiarity with the
best existing algorithms for subgraph isomorphism, and with more broadly-applicable lessons on
algorithm design from the constraint programming literature.

Another such lesson is the importance of choosing a good set of benchmark instances. As
\citet{DBLP:journals/heuristics/Gent98} observes:

@@

We hope that this paper provides some much-needed help in this direction, as some recent
evaluations work only with extremely easy instances. For example, \citet{DBLP:journals/tcbb/BonniciG17}
compared various non-constraint-based approaches to LAD using benchmark sets whose
difficulties are measured between microseconds and milliseconds. Their results suggest that LAD is
an order of magnitude slower than other approaches. This should not surprise us: the initialisation
costs of LAD are non-trivial due to it constructing domains, calculating neighbourhood degree
sequences, and performing all-different filtering. However, the risks of occasionally encountering
exponential performance from weaker algorithms on easy instances should now be clear.

@@

Unfortunately, we saw that this aspect of VF2's performance has had practical consequences, with the
most notable being the misdesign of graph database systems. By not trying even the simplest constraint
programming techniques from the literature, and disregarding all that is known about the empirical
hardness of \NP-complete problems, members of the graph databases community have reached the mistaken
conclusion that extensive research into supporting techniques is important, whereas really what
they are doing is working around some but not all of the weaknesses in their chosen algorithms for subgraph
isomorphism. With this new understanding of what does and does not make subgraph
isomorphism hard, it is time for a radical rethink of how graph database systems work.
